### 2025-04-27 analysis/1_study1.qmd
update

---
title: "Analysis - Study 1"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    toc: true
    code-fold: true
    self-contained: false
execute: 
  cache: true
---

```{r}
#| message: false
#| warning: false

options(mc.cores = parallel::detectCores() - 2)

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(ggdist)
library(brms)
library(cmdstanr)
library(rstan)

set.seed(123)
```

## Data Preprocessing

```{r}
#| eval: false

dfraw <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace/refs/heads/main/data/data.csv") |> 
  mutate(Real = (Belief_Answer + 1) / 2,  # Rescale
         Orientation = ifelse(Sexual_Orientation == "Straight ", "Heterosexual", Sexual_Orientation),
         Item = gsub(".jpg", "", Stimulus)) |> 
  filter(Sex %in% c("Male", "Female"), Orientation %in% c("Heterosexual", "Bisexual", "Homosexual")) |> 
  mutate(Relevance = case_when(
    Sex == "Male" & Orientation == "Heterosexual" & grepl("NF", Stimulus) ~ "Relevant",
    Sex == "Female" & Orientation == "Heterosexual" & grepl("NM", Stimulus) ~ "Relevant",
    Sex == "Male" & Orientation == "Bisexual" & grepl("NF", Stimulus) ~ "Relevant",
    Sex == "Female" & Orientation == "Bisexual" & grepl("NM", Stimulus) ~ "Relevant",
    Sex == "Male" & Orientation == "Homosexual" & grepl("NM", Stimulus) ~ "Relevant",
    Sex == "Female" & Orientation == "Homosexual" & grepl("NF", Stimulus) ~ "Relevant",
    .default = "Irrelevant"
  )) |> 
  mutate(SexOrientation = paste(Sex, "-", Orientation)) |>
  filter(Relevance == "Relevant", Orientation == "Heterosexual")

# Nudge mid-values to avoid model fitting issues
dfraw$Real[dfraw$Real == 0.5] <- sample(c(0.49, 0.51), sum(dfraw$Real == 0.5), replace = TRUE)  # Avoid 0.5

dfraw |> 
  select(Participant, Sex, Item, Real, Attractive, Beauty) |> 
  write.csv("../data/sample1.csv", row.names = FALSE)
```

```{r}
#| eval: false

# For demographic section
dfraw |> 
  mutate(n_stims = n(), .by = "Participant") |> 
  select(Participant, Age, Sex, n_stims) |> 
  slice(1, .by = "Participant") |> 
  report::report_participants()

report::report(dfraw$Sex)
```

```{r}
#| code-fold: false

df <- read.csv("https://raw.githubusercontent.com/RealityBending/FictionChoco/refs/heads/main/data/sample1.csv")
df$Sex <- factor(df$Sex, levels = c("Male", "Female"))
```

## Models

::: {.panel-tabset}

### ZOIB

```{r}
m1_zoib <- readRDS("models/sample1_zoib.rds")
params_table <- function(m) {
  d <- as.data.frame(m, variable = insight::find_parameters(m, effects = "fixed", flatten  = TRUE))
  t <- bayestestR::describe_posterior(d, effects = "fixed", component = "all", ci_method = "hdi", test = "p_direction")
  out <- insight::format_table(t)
  out |> 
    mutate(col = t$pd) |> 
    gt::gt() |> 
    gt::data_color(columns = "col", method = "numeric", target_columns = "pd", 
                   row = col > 0.95, palette = "yellow") |> 
    gt::data_color(columns = "col", method = "numeric", target_columns = "pd", 
                   row = col > 0.97, palette = "green") |> 
    gt::cols_hide("col")
 }
params_table(m1_zoib)
summary(m1_zoib, robust = TRUE)

make_ppcheck  <- function(m, fill = "#FF9800") {
  pred <- m |> 
    modelbased::estimate_prediction(df, keep_iterations = 100, iterations = 100) |>
    select(Sex, starts_with("iter_")) |> 
    bayestestR::reshape_iterations()

  pred |>
    ggplot(aes(x = iter_value)) +
    stat_bin(aes(group = iter_group, y = after_stat(count / max(count))), position = "identity",
             geom="bar", breaks = seq(0, 1.0, by = 0.02), pad = TRUE,
             alpha = 0.01, fill = fill) +
    stat_bin(data = df, aes(x = Real, y = after_stat(count / max(count))), 
             geom="step", breaks = seq(0, 1.0, by = 0.02), 
             linewidth = 0.5, color = "#9C27B0", position = position_nudge(x=-0.01), pad = TRUE) + 
    scale_y_sqrt(expand = c(0, 0.01)) +
    scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c("      0%", "50%", "100%      ")) +
    theme_minimal() +
    theme(axis.title.x = element_text(face = "bold"), 
          axis.title.y = element_blank(), 
          axis.text.y = element_blank(), 
          strip.background = element_rect(fill = "lightgrey", color = NA)) +
    facet_wrap(~Sex, scales = "free_y") +
    coord_cartesian(xlim = c(-0.003, 1.005), ylim=c(0, 1), expand = TRUE)
}

p1_zoib <- make_ppcheck(m1_zoib, fill = "#FF9800") +
  labs(x = "Zero-Inflated Beta (ZOIB)")
p1_zoib
```

- Women: higher real mean
- Women: no difference in precision
- Women: less extreme values BUT more ones

### Beta-Gate

```{r}
m1_betagate <- readRDS("models/sample1_betagate.rds")
params_table(m1_betagate)
summary(m1_betagate, robust = TRUE)

p1_betagate <- make_ppcheck(m1_betagate, fill = "#FF5722") +
  labs(x = "Beta-Gate")
p1_betagate

# d <- brms::as_draws(m1_betagate, draw=1:10, variable = "^b_", regex = TRUE) |> 
#   bayestestR::describe_posterior(centrality = "median", test = "p_direction")
```

- Women: higher real mean
- Women: no difference in precision
- Women: less extreme values BUT more ones

### CHOCO

```{r}
m1_choco <- readRDS("models/sample1_choco.rds")
params_table(m1_choco)
summary(m1_choco, robust = TRUE)

p1_choco <- make_ppcheck(m1_choco, fill = "#7B3F00") +
  labs(x = "CHOCO")
p1_choco
```

- Women: higher probability of "Real"
- Women: Less confidence when "Fake"
- Women: no difference in precision
- Women: less extreme values (BUT possibly more ones)

```{r}
visualize_dchoco <- function(m, newdata) {

  p <- modelbased::estimate_prediction(m, data = newdata, predict = "mu", iterations = 200)
  confright <- modelbased::estimate_prediction(m, data = newdata, predict = "confright", iterations = 200)
  confleft <- modelbased::estimate_prediction(m, data = newdata, predict = "confleft", iterations = 200)
  precright <- modelbased::estimate_prediction(m, data = newdata, predict = "precright", iterations = 200)
  precleft <- modelbased::estimate_prediction(m, data = newdata, predict = "precleft", iterations = 200)
  pex <- modelbased::estimate_prediction(m, data = newdata, predict = "pex", iterations = 200)
  bex <- modelbased::estimate_prediction(m, data = newdata, predict = "bex", iterations = 200)

  data <- data.frame()
  x <- seq(0, 1, length.out = 1001)
  for(i in 1:nrow(newdata)) {
     data <- data.frame(
       x = c(-0.005, x, 1.005),
       y = c(0, cogmod::dchoco(x, p=p$Predicted[i], confright=confright$Predicted[i], confleft = confleft$Predicted[i],
                          precright = precright$Predicted[i], precleft = precleft$Predicted[i],
                          pex = pex$Predicted[i], bex = bex$Predicted[i], pmid = 0), 0)
     ) |> 
       merge(newdata[i, ]) |> 
       rbind(data)
  }
  data
}

pred <- visualize_dchoco(m1_choco, newdata=insight::get_datagrid(m1_choco))
p1_density <- pred |> 
  mutate(y = ifelse(x %in% c(0, 1), y * 3, y)) |>  # Inflate extremes for visibility
  ggplot(aes(x=x, y=y, color=Sex)) +
  geom_line(linewidth = 1.5) +
  scale_color_manual(values = c("Male"="#2196F3", "Female"="#E91E63")) +
  scale_x_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1), 
                     labels = c("0%\nAI-Generated", "25%", "50%", "75%", "100%\nPhotograph")) +
  theme_minimal() +
  labs(y = "Distribution", x = "Do you think this image is real?") +
  theme(axis.text.y = element_blank(),
        axis.title.x = element_text(face = "bold"))
p1_density
```

:::

## Model Comparison

### Fitting Duration

```{r}
rbind(
  mutate(attributes(m1_zoib$fit)$metadata$time$chain, Model="ZOIB"),
  mutate(attributes(m1_betagate$fit)$metadata$time$chain, Model="BEXT"),
  mutate(attributes(m1_choco$fit)$metadata$time$chain, Model="CHOCO")
) |> 
  mutate(Model = fct_relevel(Model, "ZOIB", "BEXT", "CHOCO"),
         total = total / 60) |>
  ggplot(aes(x = Model, y = total, fill = Model)) +
  geom_boxplot() +
  labs(y = "Sampling Duration (min)") +
  scale_y_log10() +
  scale_fill_manual(values = c("#FF9800", "#FF5722", "#795548")) +
  theme_minimal() 
```

### Fit Quality

```{r}
loocomp <- loo::loo_compare(m1_zoib, m1_betagate, m1_choco, criterion = "waic")

parameters(loocomp, include_ENP = TRUE)

report::report(loocomp)

p1_zoib | p1_betagate | p1_choco
```

## Individual-Level Parameters

### Reliability

::: {.panel-tabset}

#### ZOIB

```{r}
table_dvour <- function(g) {
  performance::performance_dvour(g) |> 
    arrange(desc(D_vour)) |> 
    gt::gt() |> 
    gt::data_color("D_vour", palette = c("red", "red", "orange", "yellow", "green"), domain = c(0, 1)) |> 
    gt::fmt_auto()
}

g1_zoib <- modelbased::estimate_grouplevel(m1_zoib)
table_dvour(g1_zoib)
```

#### Beta-Gate

```{r}
g1_betagate <- modelbased::estimate_grouplevel(m1_betagate)
table_dvour(g1_betagate)
```

#### CHOCO

```{r}
g1_choco <- modelbased::estimate_grouplevel(m1_choco)
table_dvour(g1_choco)
```

```{r}
colors <- c("p" = "#9C27B0", "confright" = "#795548", "confleft" = "#795548",
            "precright" = "#607D8B","precleft" = "#607D8B", "pex" = "#FF9800", "bex" = "#F44336")

p2_dvour <- g1_choco |> 
  mutate(Component = ifelse(Component == "conditional", "p",  Component),
         Component = fct_relevel(Component, "p", "confright", "confleft", "precright", "precleft", "pex", "bex"),
         Group = fct_relevel(Group, "Participant", "Item"),
         .group = paste0(Group, " (", Component, ")")) |>
  ggplot(aes(x = Median, y = Level)) +
  geom_pointrange(aes(xmin = CI_low, xmax = CI_high, color= Component), linewidth = 0.3, size = 0.05) +
  facet_grid(Group~Component, scales = "free", switch="both") +
  ggside::geom_xsidedensity(aes(group = .group, fill = Component), color = NA) +
  ggside::ggside(scales  = "free") +
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  theme_minimal() +
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        strip.placement.x = "outside",
        strip.background = element_rect(fill = NA, color = NA),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        legend.position = "none")
p2_dvour
```

:::

### Correlation

```{r}
features1 <- df |> 
  summarize(Sample = "Sample1", 
            Empirical_p = sum(Real > 0.5) / n(), 
            Empirical_conf = mean(abs(Real - 0.5) * 2),
            Empirical_pex = sum(Real %in% c(0, 1)) / n(),
            Empirical_bex = sum(Real == 1) / sum(Real %in% c(0, 1)),
            Empirical_mu = mean(Real),
            .by = "Participant")

features1 <- reshape_grouplevel(g1_zoib, indices = "Median", group = "Participant") |> 
  datawizard::data_rename("Intercept", "mu") |> 
  select(-coi_Intercept) |> 
  datawizard::data_addprefix("ZOIB_", exclude = "Participant") |> 
  full_join(features1, by = "Participant")

features1 <- reshape_grouplevel(g1_betagate, indices = "Median", group = "Participant") |> 
  datawizard::data_rename("Intercept", "mu") |> 
  select(-bex_Intercept) |> 
  datawizard::data_addprefix("BGATE_", exclude = "Participant") |> 
  full_join(features1, by = "Participant")

features1 <- reshape_grouplevel(g1_choco, indices = "Median", group = "Participant") |> 
  datawizard::data_rename("Intercept", "p") |> 
  select(-bex_Intercept) |> 
  datawizard::data_addprefix("CHOCO_", exclude = "Participant") |> 
  full_join(features1, by = "Participant")
```

```{r}
cordata <- correlation::correlation(select(features1, -Participant, -Sample), p_adjust = "none", redundant = TRUE) |> 
  correlation::cor_sort() |> 
  correlation::cor_lower() |>
  as.data.frame() |> 
  mutate(Parameter2 = fct_rev(Parameter2),
         Parameter1 = fct_relabel(Parameter1, \(x) gsub("_Intercept", "", x)),
         Parameter2 = fct_relabel(Parameter2, \(x) gsub("_Intercept", "", x)),
         Parameter1 = fct_relabel(Parameter1, \(x) gsub("_", " - ", x)),
         Parameter2 = fct_relabel(Parameter2, \(x) gsub("_", " - ", x)))

p2_cor <- cordata |> 
  ggplot(aes(x=Parameter2, y = Parameter1)) +
  geom_tile(aes(fill = r)) +
  geom_text(aes(label = insight::format_value(r, 2, zap_small = TRUE, lead_zero = FALSE)), color = ifelse(cordata$p < 0.01, "black", "grey")) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    legend.position = "none"
  )
p2_cor
```

Choco parameters are the most interpretable (relate to empirical indices).

## Figures

```{r}
#| eval: true

p1_a <- wrap_elements((p1_zoib | p1_betagate | p1_choco) + 
         patchwork::plot_annotation(title = "Posterior Predictive Checks", theme = theme(plot.title = element_text(face = "bold"))))

p1_b <- wrap_elements(p1_density + 
         patchwork::plot_annotation(title = "Effect of Sex", theme = theme(plot.title = element_text(face = "bold"))))

p1 <- p1_a / p1_b

p2_a <- wrap_elements(p2_dvour + 
         patchwork::plot_annotation(title = "Reliability of CHOCO Participant-Level Estimates", theme = theme(plot.title = element_text(face = "bold"))))

p2_b <- wrap_elements(p2_cor + 
         patchwork::plot_annotation(title = "Correlation of Participant-Level Estimates", theme = theme(plot.title = element_text(face = "bold"))))
p2 <- p2_a / p2_b + plot_layout(heights = c(1.5, 1))

ggsave("../paper/figures/fig3.png", p1, width = 8, height = 8, dpi = 300)
ggsave("../paper/figures/fig4.png", p2, width = 8, height = 12, dpi = 300)
```