### 2025-04-27 paper/manuscript.tex

\documentclass[
  jou,
  floatsintext,
  longtable,
  nolmodern,
  notxfonts,
  notimes,
  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}

\usepackage{amsmath}
\usepackage{amssymb}

\RequirePackage{longtable}
\RequirePackage{threeparttablex}

\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-.5em}%
	{\normalfont\normalsize\bfseries\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{0.5em}%
	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
	{-\z@\relax}%
	{\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

\usepackage{longtable, booktabs, multirow, multicol, colortbl, hhline, caption, array, float, xpatch}
\usepackage{subcaption}
\renewcommand\thesubfigure{\Alph{subfigure}}
\setcounter{topnumber}{2}
\setcounter{bottomnumber}{2}
\setcounter{totalnumber}{4}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.85}
\renewcommand{\textfraction}{0.15}
\renewcommand{\floatpagefraction}{0.7}

\usepackage{tcolorbox}
\tcbuselibrary{listings,theorems, breakable, skins}
\usepackage{fontawesome5}

\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{ACACAC}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582EC}
\definecolor{quarto-callout-important-color-frame}{HTML}{D9534F}
\definecolor{quarto-callout-warning-color-frame}{HTML}{F0AD4E}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02B875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{FD7E14}

%\newlength\Oldarrayrulewidth
%\newlength\Oldtabcolsep

\usepackage{hyperref}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{newtx}

\defaultfontfeatures{Scale=MatchLowercase}
\defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}

\title{\textbf{Introducing the Choice-Confidence (CHOCO) Model for
Bimodal Scales Data: Application to the Relationship between Reality
Beliefs about AI-Generated Faces and Beauty}}

\shorttitle{Choice-Confidence (CHOCO) Model}

\usepackage{etoolbox}

\authorsnames[{1,2},{1},{1}]{Dominique Makowski,Ana Neves,Andy Field}

\authorsaffiliations{
{School of Psychology, University of Sussex},{Sussex Centre for
Consciousness Science, University of Sussex}}

\leftheader{Makowski, Neves and Field}

\abstract{TO DO. }

\keywords{subjective scales, choice confidence model, reality beliefs,
AI-generated faces, attractiveness}

\authornote{\par{\addORCIDlink{Dominique
Makowski}{0000-0001-5375-9967}}\par{\addORCIDlink{Ana
Neves}{0009-0006-0020-7599}}\par{\addORCIDlink{Andy Field}{true}} 
\par{ }
\par{     \begin{tcolorbox}[enhanced jigsaw, left=2mm, arc=.35mm, bottomrule=.15mm, colframe=quarto-callout-note-color-frame, leftrule=.75mm, breakable, rightrule=.15mm, toprule=.15mm, colback=white, opacityback=0]

This preprint is a non-peer-reviewed work from the
\href{https://realitybending.github.io/}{\textbf{Reality Bending Lab}}.
\begin{center}
\includegraphics[width=0.2\textwidth,height=\textheight]{manuscript_files/mediabag/ReBeL_LogoOnly_hu114.png}
\end{center}

\end{tcolorbox}  Author roles were classified using the Contributor Role
Taxonomy (CRediT; https://credit.niso.org/) as follows:  Dominique
Makowski:   Conceptualization, Data curation, Formal Analysis, Funding
acquisition, Investigation, Methodology, Project
administration, Resources, Software, Supervision, Validation, Visualization, Writing
-- original draft; Ana Neves:   Data curation, Writing -- original
draft, Writing -- review \& editing; Andy Field:   Writing -- original
draft, Writing -- review \& editing}
\par{Correspondence concerning this article should be addressed
to Dominique
Makowski, Email: \href{mailto:D.Makowski@sussex.ac.uk}{D.Makowski@sussex.ac.uk}}
}

\usepackage{pbalance} 
\usepackage{float}
\makeatletter
\let\oldtpt\ThreePartTable
\let\endoldtpt\endThreePartTable
\def\ThreePartTable{\@ifnextchar[\ThreePartTable@i \ThreePartTable@ii}
\def\ThreePartTable@i[#1]{\begin{figure}[!htbp]
\onecolumn
\begin{minipage}{0.5\textwidth}
\oldtpt[#1]
}
\def\ThreePartTable@ii{\begin{figure}[!htbp]
\onecolumn
\begin{minipage}{0.5\textwidth}
\oldtpt
}
\def\endThreePartTable{
\endoldtpt
\end{minipage}
\twocolumn
\end{figure}}
\makeatother

\makeatletter
\let\endoldlt\endlongtable		
\def\endlongtable{
\hline
\endoldlt}
\makeatother

\newenvironment{twocolumntable}% environment name
{% begin code
\begin{table*}[!htbp]%
\onecolumn%
}%
{%
\twocolumn%
\end{table*}%
}% end code

\urlstyle{same}

\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

% From https://tex.stackexchange.com/a/645996/211326
%%% apa7 doesn't want to add appendix section titles in the toc
%%% let's make it do it
\makeatletter
\xpatchcmd{\appendix}
  {\par}
  {\addcontentsline{toc}{section}{\@currentlabelname}\par}
  {}{}
\makeatother

%% Disable longtable counter
%% https://tex.stackexchange.com/a/248395/211326

\usepackage{etoolbox}

\makeatletter
\patchcmd{\LT@caption}
  {\bgroup}
  {\bgroup\global\LTpatch@captiontrue}
  {}{}
\patchcmd{\longtable}
  {\par}
  {\par\global\LTpatch@captionfalse}
  {}{}
\apptocmd{\endlongtable}
  {\ifLTpatch@caption\else\addtocounter{table}{-1}\fi}
  {}{}
\newif\ifLTpatch@caption
\makeatother

\begin{document}

\maketitle

\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\setlength\LTleft{0pt}

\section{Introduction}\label{introduction}

Despite significant advancements in psychological science following the
replication crisis (\citeproc{ref-open2015estimating}{Collaboration,
2015}), its progress is still hindered by its sub-optimal (or
inappropriate) usage of statistical tools
(\citeproc{ref-makowski2023we}{Makowski \& Waggoner, 2023}). A prevalent
issue is the continued reliance on linear models that assume normally
distributed (Gaussian) data - as this assumption often does not hold
true for many types of psychological outcomes. For instance, reaction
times typically exhibit skewed distributions, choices can be represented
as binary variables, and count data consists of strictly positive
integers. Applying models that presume normality and model the ``mean''
of the outcome variable can lead to misinterpretation and potentially
misleading conclusions when applied indiscriminately. It is thus
important that psychologists use models that can best describe the data
they collect, to fully exploit them and bring more nuance and accuracy
to their conclusions.

Among the most commonly collected data in psychology are responses on
subjective scales, such as Likert-type items or visual analog scales,
which exhibit some fundamental properties: these responses are bounded
(and can be rescaled to a 0-1 range) and frequently display clustering
at the extremes. Traditional linear models being ill-suited for such
data, researchers have turned to using Beta distributions to model this
data (instead of Gaussian), suited for continuous data within the (0,1)
interval (i.e., excluding extreme responses). To address the frequent
occurrence of exact zeros and ones (i.e., extreme values), zero-one
inflated beta (ZOIB) models have been developed
(\citeproc{ref-ospina2012general}{Ospina \& Ferrari, 2012}) to
accommodate the excess of boundary values by incorporating additional
components that model the probabilities of responses at 0 and 1 as a
separate, independent process.

\subsection{The Beta-Gate Model}\label{the-beta-gate-model}

The Beta-Gate model is a reparametrized Ordered Beta model
(\citeproc{ref-kubinec2023ordered}{Kubinec, 2023})\footnote{In the
  Ordered Beta model, the cutpoints on the log-scale are directly used
  as parameters, instead of being derived from \emph{pex} and
  \emph{bex}.} available in the \emph{cogmod} package in R
(https://github.com/DominiqueMakowski/cogmod), in which participants'
answers on bounded scales are conceptualized as latent responses that
can fall past a pair of probabilistic ``gates'' (or cutpoints) that
control whether the response is recorded as an extreme (0 or 1) or as a
nuanced, continuous value in between (Figure~\ref{fig-one}). These
distance of these gates from the edges of the scale varies based on two
interpretable parameters: \emph{pex} (the propensity by which people are
likely to answer extreme values), and \emph{bex}, a bias toward the
upper extreme (1) versus the lower (0). A person's internal response
that lies close to the edge might be ``caught'' by a gate and recorded
as an extreme, while others pass through to express a continuous
response (Beta-distributed with \(\mu\) (\emph{mu}) and \(\phi\)
(\emph{phi}) as its mean and precision parameters). The Beta-Gate model
is based on the idea that extreme values can emerge not just from a
fundamentally different underlying processes - as assumed in ZOIB models
- but from a common process governed by thresholds of decisiveness and
confidence.

\begin{figure}

\caption{\label{fig-one}The Beta-Gate Distribution is a reparametrized
ordered Beta model (Kubinec, 2023) that is governed by 4 parameters.
`Mu' and `phi' correspond to the mean and precision of the continuous
part of the distribution (between 0 and 1), and `pex' (propensity of
extremes) and `bex' (balance of extremes) indirectly control the
proportion of zeros and ones by specifying the location of the
``gates'', past which the latent response process is likely to generate
extreme values. Specifically, `pex' defines the total distance of both
gates from the extremes (in yellow), and `bex' determines the proportion
of the right gate distance relative to the left. In this example, the
total distance from the extremes is `pex' = 0.1, with 40\% (`bex' = 0.4)
of that distance being on the right (and 60\% on the left). The left
gate is thus located at 0.6, and the right at 1-0.04 = 0.96.}

\centering{

\includegraphics[width=0.5\textwidth,height=\textheight]{./figures/fig1.png}

}

\end{figure}%

Mathematically, the Beta-Gate distribution defines the observed outcome
\(x \in [0, 1]\) as a mixture of three components; a point mass at 0, a
point mass at 1, and a continuous Beta density over \(x \in (0,1)\),
scaled by the remaining probability mass. The probability of these
components are:

\begin{itemize}
\tightlist
\item
  \(P(x = 0) = \text{logistic}\left( \text{logit}(pex \cdot (1 - bex)) - \text{logit}(\mu) \right)\)
\item
  \(P(x = 1) = 1 - \text{logistic}\left( \text{logit}(1 - pex \cdot bex) - \text{logit}(\mu) \right)\)
\item
  \(P(x \in (0,1)) = 1 - P(x = 0) - P(x = 1)\)
\end{itemize}

The continuous part follows a Beta distribution with
parameters\footnote{Note that \emph{phi} is scaled in Beta-Gate models
  relative to the traditional mu/phi Beta particularization so that a
  phi of 1 corresponds to a uniform distribution - to facilitate setting
  priors on this parameter}:

\[
Beta(\alpha = \mu \cdot 2 \phi, \quad \beta = (1 - \mu) \cdot 2 \phi)
\]

\subsection{The Choice-Confidence (CHOCO)
Model}\label{the-choice-confidence-choco-model}

Decision-making is often conceptualized as involving distinct processes:
the choice itself, and the confidence associated with that choice. In
experimental paradigms, these can be somewhat disentangled by prompting
participants to make a discrete choice selection (e.g., ``True''
vs.~``False''), followed by a separate confidence rating. However, this
artificial separation makes its joint analysis difficulty, and may not
reflect real-world decision-making, where individuals often express both
choice and confidence simultaneously using a single, continuous scale.
In such scales, each side can represent a distinct latent category, and
the distance from the midpoint can indicates the level of confidence or
certainty. This integrated response format typically results in bimodal
distributions, with peaks corresponding to the mean confidence on either
side. Traditional beta regression models, which assume unimodal
distributions within the (0,1) interval, are ill-suited for such data.
One alternative is to transform the data into two variables a
posteriori: binarizing the side to represent choice and calculating the
absolute distance from the midpoint to represent confidence. These can
then be modeled separately, for instance, using logistic regression for
choice and beta regression for confidence (see
\citeproc{ref-makowski2025too}{Makowski et al., 2025} for an example).
While this approach can provide additional insights into underlying
mechanisms compared to a unique model, it assumes psychological and
statistical independence between choice and confidence, which may not
hold true in practice.

\begin{figure}

\caption{\label{fig-two}The CHOCO Model uses a mixture of Beta-Gate
distributions to model separately the right and left sides of the scale
(e.g., a rating of whether a statement was `Truth' vs.~`Lie'), as well
as their relative proportion. In this example, the participants are more
likely overall (p = 60\%) to select the right side of the scale (`Lie')
than the left (`Truth'). They are also more confident in their choice
(confright = 0.6 vs.~confleft = 0.3). Extreme values (zeros and ones)
are governed by the same mechanism as for Beta-Gate models.}

\centering{

\includegraphics[width=0.5\textwidth,height=\textheight]{./figures/fig2.png}

}

\end{figure}%

To model data of subjective scales in which the left and right sides can
be conceptualized as two different choices (e.g., True/False,
Agree/Disagree, etc.) and the magnitude of the response (how much the
cursor is set away from the mid-point) as the confidence, we introduce
the Choice-Confidence (CHOCO) model (Figure~\ref{fig-two}). It consists
of a three‐part mixture on \(x \in [0, 1]\):

\begin{itemize}
\tightlist
\item
  An (optional) point-mass at the midpoint \emph{mid} (typically 0.5) of
  weight \(p_{\text{mid}}\) for undecided or neutral responses.\\
\item
  A left-choice component governed by a Beta-Gate density on the
  rescaled variable \(x/mid\) with mean \emph{1 - confleft}, precision
  \emph{precleft}, and boundary-excess parameter \emph{pex(1 - bex)}.\\
\item
  A right-choice component governed by a Beta-Gate density on the
  rescaled variable \((x - mid)/(1 - mid)\) with mean \emph{confright},
  precision \emph{precright}, and boundary-excess parameter \emph{pex x
  bex}.
\end{itemize}

The overall probability of the right choice (relative to the left
choice) is controlled by a main parameter \emph{p}. The full CHOCO
density is:

\[
\small
\begin{cases}
pmid, & x = mid, \\[6pt]
(1 - pmid)(1 - p)\,\dfrac{1}{mid} \cdot \mathrm{BetaGate}\left(\dfrac{x}{mid}\right), & 0 < x < mid, \\[6pt]
(1 - pmid)\,p\,\dfrac{1}{1 - mid} \cdot \mathrm{BetaGate}\left(\dfrac{x - mid}{1 - mid}\right), & mid < x < 1. \\
\end{cases}
\]

By coupling choice probability \(p\), midpoint mass \(p_{\text{mid}}\),
and side‐specific Beta-Gate parameters (\(conf\), \(prec\), \(pex\),
\(bex\)), CHOCO flexibly captures both bimodality and confidence
intensity in a single unified model. Despite this theoretical appeal, it
is unclear whether this heavily parametrized model can be estimated
reliably from data, and whether it can provide more useful insights than
simpler alternatives.

\subsection{Aim of the Present Study}\label{aim-of-the-present-study}

\textbf{Study 1} aims to evaluate the CHOCO model's ability to better
capture subjective scale responses that (potentially) reflect an
underlying discrete choice, in comparison to existing models such as the
ZOIB and Beta-Gate. Specifically, we will assess whether 1) CHOCO
provides improved model fit, 2) yields deeper insights into
population-level effects than traditional approaches (gender differences
in reality beliefs), and 3) allows for the reliable estimation of
interpretable individual-level parameters through random effects.

To this end, we analyze data from two separate studies in which
participants judged whether a face image was AI-generated (``fake'') or
a real photograph. \textbf{Study 2} will apply this model to more subtle
effects, such as the effect perceived facial attractiveness and beauty
on reality judgments.

\section{Study 1}\label{study-1}

\textbf{TODO}: Write some introduction about reality beliefs in the
context of ambiguity. Underline that what drives judgments and beliefs
of reality remains unclear. And this is important in today's post-truth
era.

Another unknown is whether these beliefs are fully stimuli-driven or
whether there are some stable individual tendencies that are at stake.

\subsection{Methods}\label{methods}

\subsubsection{Participants}\label{participants}

The sample is the same as in (\citeproc{ref-makowski2025too}{Makowski et
al., 2025}). We included all heterosexual male and female participants,
for a final sample of 125 participants (Mean age = 28.6, SD = 9.2,
range: {[}19, 66{]}; Sex: 44.8\% females). We included only sexually
relevant stimuli for men and women (i.e., all 89 female faces for men
and 20 male faces for women).

\subsubsection{Procedure}\label{procedure}

In the first phase, participants viewed 109 neutral-expression
photographs of faces (random order, display time of 500 ms) from the
American Multiracial Face Database
(\citeproc{ref-chen2021broadening}{Chen et al., 2021}). After each
image, participants rated the face on trustworthiness, familiarity,
attractiveness, and beauty using visual analog scales. In the second
phase, participants were informed that ``about half of the previously
seen images were AI-generated''. The same faces were presented again in
a new random order (same display time), followed by ratings of
``reality'' (whether they believed the image was AI-generated - left
anchor - or a real photograph - right anchor). Makowski et al.
(\citeproc{ref-makowski2025too}{2025}) showed that most stimuli yielded
highly variable reality ratings, suggesting that idiosyncratic cognitive
processes were at play.

\subsubsection{Data Analysis}\label{data-analysis}

We fitted 3 models to predict the reality ratings: a ZOIB model, a
Beta-Gate model, and the CHOCO model. For all models and each parameter,
the full formula was entered:
\(Real\sim Sex + (1 | Participant) + (1 | Item)\) (with Sex as the main
predictor and participants and items entered as random intercepts). The
models were run using \emph{brms}
(\citeproc{ref-burkner2017brms}{Bürkner, 2017}) R package, and analyzed
using the \emph{easystats} collection of packages {[}\textbf{REFS}{]}.
To maximize the comparability across models We used the default priors
(uniform) for all models, and we ran 16 chains of 1400 iterations each
on the University of Sussex High-Performance Computing (HPC) cluster.

Model comparisons were performed using the \emph{loo} R package
(\citeproc{ref-vehtari2017practical}{Vehtari et al., 2017}), which
computes the Widely Applicable Information Criterion (WAIC) and
estimates the Expected Log Predictive Density (ELPD) and penalizes the
number of parameters. We assessed model performance by examining ELPD
differences and their standard errors (SE), reporting corresponding
\emph{p}-values to determine significant differences in predictive
accuracy.

For the population-level effects, we will consider significant and
report (using the median of the posterior distribution) effects for
which the 95\% Credible Interval (CI) does not include zero {[}and when
the probability of direction \emph{pd} is \textgreater{}
\textasciitilde97\%; Makowski et al.
(\citeproc{ref-makowski2019indices}{2019}){]}. For the individual-level
parameters (i.e., the random intercepts of each parameter for each
participant and each item), we will first analyze their reliability
using the Variance-Over-Uncertainty Ratio index (\emph{D-vour}). This
index, implemented in the \emph{performance} package
(\citeproc{ref-ludecke2021performance}{Lüdecke et al., 2021}), is
inspired by recent work on mixed models reliability
(\citeproc{ref-rouder2024hierarchical}{Rouder \& Mehrvarz, 2024};
\citeproc{ref-williams2021beneath}{Williams et al., 2021}), and
corresponds to the normalized ratio of observed variability to
uncertainty in random effect estimates, defined as:

\[
D_{\text{vour}} = \frac{\sigma_B^2}{\sigma_B^2 + \mu_{\text{SE}}^2}
\]

Where \(\sigma_B^2\) is the between-group variability (computed as the
SD of the random effect point-estimates) and \(\mu_{\text{SE}}^2\) is
the mean squared uncertainty in random effect estimates (i.e., the
average uncertainty). We use as \emph{D-vour} = 0.666 as the threshold
for moderately reliable random effect estimates, which corresponds to a
2:1 ratio of between-group variance to uncertainty.

Finally, we will run a correlation analysis of the models'
individual-level estimates against ``empirical'' (indices computed
directly on the observed data), including the empirical \emph{p}
(\(P(y > 0.5)\)), the overall \emph{conf} (\(mean(|y - 0.5|)\)),
\emph{pex} (\(P(y \in [0, 1])\)) and \emph{bex}
(\(P(y == 1) / P(y \in [0, 1])\)), assessing whether the model's
estimate are in-line with easily interpretable indices.

\subsection{Results}\label{results}

The reproducible code and full result report are available at
https://github.com/RealityBending/FictionChoco.

\subsubsection{Model Comparison}\label{model-comparison}

\begin{figure*}[!htbp]

{\caption{{TODO.}{\label{fig-three}}}}

\includegraphics[width=1\textwidth,height=\textheight]{./figures/fig3.png}

\end{figure*}

The models did converge without divergent transitions, and the effective
sample size was sufficient for all parameters (all
\(n_{\text{eff}} > 1000\)). The difference in predictive accuracy, as
indexed by Expected Log Predictive Density (ELPD-WAIC), suggests that
the \emph{CHOCO} is the best model (\(ELPD = -138.45\)), followed by
\emph{Beta-Gate} (\(\Delta_{ELPD} = -1875.31 \pm 61.18, p < .001)\) and
\emph{ZOIB} (\(\Delta_{ELPD} = -1916.11 \pm 61.75, p < .001\)). See
Figure~\ref{fig-three} for the posterior predictive checks, showing that
only the CHOCO model managed to capture the bimodal distribution of
data.

\subsubsection{Effect of Sex}\label{effect-of-sex}

The ZOIB model suggested that women had higher mean (\emph{mu}
parameter) scores of reality beliefs
(\(\Delta_{\text{Female}} = 0.16,~95\%~ CI~[0.01, 0.32], pd = 98.10\%\)),
less extreme values (\emph{zoi} parameter; STATS) but more ones relative
to zeros (\emph{coi} parameter; STATS).

The Beta-Gate model similarly suggested that women had higher mean
(\emph{mu} parameter) scores of reality beliefs (STATS), less extreme
values (\emph{pex} parameter; STATS), and a greater tendency to answer
one relative to zero (\emph{bex} parameter; STATS).

The CHOCO model shows that women had a higher probability \emph{p} of
judging faces as real (STATS), but are not more confident when doing so
(\emph{confright} parameter; STATS). However, they were less confident
when answering that an image was AI-generated (\emph{confleft}
parameter; STATS). There were also less likely to produce extreme
answers (\emph{pex} parameter; STATS), but no strong evidence supporting
a directional bias was observed (\emph{bex} parameter; STATS).

Across all these models, no effect of Sex on the prediction parameter
was observed.

\subsubsection{Individual-Level
Parameters}\label{individual-level-parameters}

\begin{figure*}[!htbp]

{\caption{{TODO.}{\label{fig-four}}}}

\includegraphics[width=1\textwidth,height=\textheight]{./figures/fig4.png}

\end{figure*}

The ZOIB model estimated reliable variability in the participant's
\emph{phi} parameter (D-vour = 0.88) and \emph{zoi} parameter (D-vour =
0.84), as well as in the \emph{mu} parameter related to individual items
(D-vour = 0.82). Moderate reliability was also observed for items in the
\emph{coi} parameter (D-vour = 0.71) and for participants in the
\emph{mu} parameter (D-vour = 0.69). The Beta-Gate model yielded similar
results: a high reliability of participant's \emph{phi} parameter
(D-vour = 0.88), \emph{pex} parameter (D-vour = 0.84). The \emph{mu}
parameter's variability was reliably captured for items (D-vour = 0.85)
and moderately for participants (D-vour = 0.72). Of note that the
parameters of Beta-Gate and ZOIB correlate almost perfectly, underlining
their empirical similarity despite a different parametrization.

The CHOCO model yielded reliable estimates for all parameters except
\emph{bex} for participants (\emph{confright} D-vour = 0.94,
\emph{confleft} D-vour = 0.91, \emph{pex} D-vour = 0.80, \emph{p} D-vour
= 0.73, \emph{precright} D-vour = 0.79, \emph{precleft} D-vour = 0.73).
Item's variability was primarily reflected through the \emph{p}
parameter (D-vour = 0.86).

Finally, the empirical average correlated the most strongly with CHOCO's
\emph{p} (r = .94), rather than ZOIB's \emph{mu} (r = .74) or
Beta-Gate's \emph{mu} (r = .82). The empirical overall confidence was
the strongest correlated with CHOCO's \emph{confright} (r = .91),
followed by ZOIB's \emph{phi} (r = -.86), Beta-Gate's phi (r = -.83),
and other CHOCO's parameters. The empirical proportion of ``right''
answer \emph{p} correlated the strongest with CHOCO's \emph{p} (r =
0.94), followed by Beta-Gate's \emph{mu} (r = .82) and ZOIB's \emph{mu}
(r = .77). The empirical \emph{pex} correlated the strongest with ZOIB's
\emph{zoi} (r = .90), and Beta-Gate's \emph{pex} (r = .90), and CHOCO's
\emph{pex} (r = .88).

\subsection{Discussion}\label{discussion}

Study 1 revealed that the Choice-Confidence (CHOCO) model was a much
better fit for bimodal bounded data, compared to other alternatives like
the Zero-and-One Inflated Beta (ZOIB) and Beta-Gate (Ordered Beta)
models. It also allowed for a deeper understanding through its
interpretable parameters, offering insights into possibly distinct
cognitive mechanisms, such as the probability of answering real
vs.~fake, and the associated confidence in these two choices. This was
illustrated by modelling the effect of sex on all the CHOCO parameters.

Note that the observed gender differences are primarily presented as a
proof-of-principle, to showcase the model's ability to capture
group-level effects and to provide deeper insights compared to other
models. However, given that they were based on different items (female
and male faces), these differences might just be a reflection of stimuli
characteristics rather than true sex dymorphism in the formation of
reality beliefs.

Finally, we also show that the CHOCO model was able to capture reliable
and interpretable individual-level parameters, supporting its value to
measure inter-individual differences. An interesting dissociation
emerged between participant- and item-level variability: the latter
seemed mostly to be represented in the \emph{p} parameter, while
participants reliably varied in most of the components (aside from
\emph{bex}). This could be interpreted as that external item
characteristics primarily influence the probability of being judged as
real vs.~fake, while the expressed confidence is first and foremost an
individual characteristic.

\section{Study 2}\label{study-2}

The explosion of accessibility of state-of-the-art AI tools has made it
effortless to generate realistic images, including human faces that are
often indistinguishable from real ones \textbf{{[}REF{]}}. These
synthetic visuals are flooding the cyberspace across various domains,
such as art, advertising and entertainment, to education and
information. This technological advancement carries an important
potential for misuse, such as in disinformation campaigns, scams (e.g.,
AI-bots, identity theft), and abuse. The democratization of such
technology raises pressing concerns about the value of authenticity and
the potential erosion of media trust in our increasingly
\emph{post-truth} society.

In this evolving landscape, understanding the cognitive mechanisms that
underpin our judgments of reality becomes paramount. Despite the
increasing prevalence of digitally altered or AI-generated content,
humans still rely on certain heuristics to assess authenticity. One such
heuristic might be facial attractiveness. Attractiveness appraisals are
known to be automatic and unconscious (\citeproc{ref-hou2023neural}{Hou
et al., 2023}; \citeproc{ref-hung2016unconscious}{Hung et al., 2016};
\citeproc{ref-luo2019robust}{Luo et al., 2019}), and carry strong
real-life consequences, as demonstrated by the large body of literature
on the ``beauty premium'' (\citeproc{ref-gulati2024beautiful}{Gulati et
al., 2024}, \textbf{MORE}; \citeproc{ref-little2021facial}{Little,
2021})\footnote{As an eloquent example, Monk Jr et al.
  (\citeproc{ref-monk2021beholding}{2021}) reported that the the
  magnitude of the earnings disparities among white women along the
  perceived attractiveness continuum exceeds in magnitude the canonical
  black-white race gap in a large representative US sample.}

\textbf{TODO:} However, it is important to allow for the possibility of
a U-shaped relationship. For instance, Peng et al.~(2020) reported that
both attractive and unattractive people sold more (underlining a
``plainness penalty'').

Although its role as a potential modulator of reality beliefs remains
underexplored, Makowski et al\ldots{} \textbf{TODO: detail findings of
that study + other studies}.

In this study, we will apply the CHOCO model to the data of MAKOWSKI,
and see whether different analysis approaches lead to different
conclusions. We will then try to replicate these findings in a new
sample of participants.

\subsection{Methods}\label{methods-1}

\subsubsection{Participants}\label{participants-1}

The first sample includes the same participants as study 1 (see above).
\textbf{TODO} for sample 2.

\subsubsection{Procedure}\label{procedure-1}

For the second sample, the procedure was relatively similar with a few
key differences. \textbf{TODO} Describe fake Face 2. The main difference
is that the subjective ratings were done on a X-point Likert scale,
rather than analog, with a mid-point value.

\subsubsection{Data Analysis}\label{data-analysis-1}

To compare the benefits of CHOCO models to a ``traditional'' analytic
approach, we started by fitting a frequentist linear mixed model to
predict reality beliefs with the formula
\(Real\sim Sex / poly(Attractive, 2) + (poly(Attractive, 2)|Participant) + (1|Item)\).
The 2nd degree raw polynomial term was included to allow for potentially
non-linear relationships (note that the first and second degree effects
of \emph{raw} polynomials can be interpreted independently as the linear
part and the ``curvy'' part of the relationship). For the CHOCO model,
mildly informative and effect-agnostic (i.e., centered at zero) priors
were used. The same formula was used for all parameters, except that
(based on the reliability analysis of Study 1), items were only included
as random effects for \emph{p}, and participants were not included for
\emph{bex}.

\textbf{TODO}. For sample 2, the analysis was based on that of Sample 1.
The main differences are 1) the inclusion of the ``Condition'' (whether
the picture was presented as ``Real'' of ``Fake'' in the first phase of
the experiment) as an additional predictor (it was entered as the only
random slope for all participant random effects); 2) the inclusion of an
additional parameters, \emph{pmid}, modelling the probability of
answering the middle-point of the Likert scale (representing an
``undecided'' option).

The same analysis was applied for Beauty.

\subsection{Results}\label{results-1}

As the complete model parameters tables are available at
https://github.com/RealityBending/FictionChoco, we will focus on
reporting noteworthy findings below.

In sample 1, the traditional approach suggested a significant linear
relationship between the mean level of ``Reality'' and attractiveness
(\(\beta_{poly1}=3.63, ~95\%~CI [2.71, 4.56], pd = ...\%\)) for males
only. The CHOCO model revealed that attractiveness had a significantly
positive linear relationship with the probability \emph{p} of judging
faces as real (STATS), but a quadratic relationship with the confidence
in real judgments (\emph{confright}; STATS). Attractiveness was also
associated with less confidence in fake judgments (\emph{confleft};
STATS), and related linearly with a stronger bias towards extreme Real
responses (\emph{bex}; STATS). These effects were present for males
only, and no relationship was found for females. The reliability of the
effect of attractiveness (the random slopes) was very low for all
parameters (D-vour \textless{} 0.1).

The effect of Beauty was the same as above, with some additional
evidence in favour of an effect in women, in particular for a linear
relationship with \emph{p} for females as well
(\(\beta_{poly1}=7.02, ~95\%~CI [-0.90, 14.97], pd = 95.91\%\)), as well
as a quadradic relationship with the confidence in fake judgments
(\emph{confleft}; STATS).

In Sample 2, the traditional approach suggested a significant linear
relationship between the mean level of ``Reality'' and attractiveness
for males (STATS) and women (STATS), with no effect of the Condition.
The CHOCO model revealed that attractiveness had a significantly
positive linear relationship with the probability \emph{p} of judging
faces as real for males (STATS) and females (STATS). It also had a
linear relationship with the confidence in real judgments
(\emph{confright}) for males (STATS) and females (STATS), as well as a
significant quadratic curvature for males only (STATS). Attractiveness
also linearly decreased the confidence in fake judgments only for males
(\emph{confleft}; STATS), but increased the probability of an undecided
response (\emph{pmid}) in women (STATS), and displayed in females only a
quadratic relationship with the amount of extreme answers (\emph{pex};
STATS) and its bias in favour of right-side extreme response
(\emph{bex}; STATS). No consistent and reliable effect of condition was
found.

The effect of Beauty was\ldots{}

\subsection{Discussion}\label{discussion-1}

The fact that the effect was less obvious in women in the sample 1 might
be explained with power (less relevant items were included for women).

\section{General Discussion}\label{general-discussion}

Little difference between Beauty and Attractiveness, aside that Beauty
revealed more effects in women. Why?

Likert-scales (with a limited number of discrete options) are commonly
used in psychology, and often modeled as continuous. While this fact has
spurred its own polarized debate (REFS), our position is that the model
used should ideally reflect the data \emph{generating} mechanism rather
than necessarily focus on the \emph{observed} data\footnote{Ideally, the
  data recording method should be aligned with the assumed generating
  mechanism, which is an experiment design issue rather than a
  statistical one.} - although the former cannot always be easily
inferred or known. For the data of Sample 2, we held the assumption that
the underlying latent distribution was CHOCO distributed (as there is no
reason to assume it would be different from Sample 1), treated as
continuous CHOCO-distributed data, despite the fact that it was
collected on a 7-point Likert scale. The motivation was to see if the
model could extrapolate and reliably estimate all the parameters based
on a limited number of data points but means in our case that,
effectively, each Beta distribution was extrapolated from the frequency
of only two unique values (or 3 for the no-extreme version). While this
is far from ideal, the CHOCO model still provided a better fit to the
data than alternative approaches. \textbf{TODO} However, it might have
lead to a lesser ability to detect finer grain changes, including
U-shaped relationships. Having 7-points (or 6 with no mid-value) is the
strict minimum to use the CHOCO model.

Future studies should assess the ability of CHOCO models to also fit
unimodal distributions (so it can be used more generally), as well as
the benefits of more parsimonious parametrization (for instance, a
unique precision parameter controlling both the left and right sides, or
a linked confidence parameter e.g.~where \emph{confleft} is expressed as
a function of \emph{confright}).

\section{Data Availability}\label{data-availability}

Data, code and everything is available at
https://github.com/RealityBending/FictionChoco. The CHOCO model is
implemented in the \emph{cogmod} R package
(https://github.com/DominiqueMakowski/cogmod).

\section{Acknowledgements}\label{acknowledgements}

We would like to thank the dissertation students from the University of
Sussex for their help in data collection.

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-burkner2017brms}
Bürkner, P.-C. (2017). Brms: An r package for bayesian multilevel models
using stan. \emph{Journal of Statistical Software}, \emph{80}, 1--28.

\bibitem[\citeproctext]{ref-chen2021broadening}
Chen, J. M., Norman, J. B., \& Nam, Y. (2021). Broadening the stimulus
set: Introducing the american multiracial faces database. \emph{Behavior
Research Methods}, \emph{53}, 371--389.

\bibitem[\citeproctext]{ref-open2015estimating}
Collaboration, O. S. (2015). Estimating the reproducibility of
psychological science. \emph{Science}, \emph{349}(6251), aac4716.

\bibitem[\citeproctext]{ref-gulati2024beautiful}
Gulati, A., Martı́nez-Garcia, M., Fernández, D., Lozano, M. A., Lepri,
B., \& Oliver, N. (2024). What is beautiful is still good: The
attractiveness halo effect in the era of beauty filters. \emph{Royal
Society Open Science}, \emph{11}(11), 240882.

\bibitem[\citeproctext]{ref-hou2023neural}
Hou, X., Shang, J., \& Tong, S. (2023). Neural mechanisms of the
conscious and subliminal processing of facial attractiveness.
\emph{Brain Sciences}, \emph{13}(6), 855.

\bibitem[\citeproctext]{ref-hung2016unconscious}
Hung, S.-M., Nieh, C.-H., \& Hsieh, P.-J. (2016). Unconscious processing
of facial attractiveness: Invisible attractive faces orient visual
attention. \emph{Scientific Reports}, \emph{6}(1), 37117.

\bibitem[\citeproctext]{ref-kubinec2023ordered}
Kubinec, R. (2023). Ordered beta regression: A parsimonious,
well-fitting model for continuous data with lower and upper bounds.
\emph{Political Analysis}, \emph{31}(4), 519--536.

\bibitem[\citeproctext]{ref-little2021facial}
Little, A. C. (2021). Facial attractiveness. \emph{Encyclopedia of
Evolutionary Psychological Science}, 2887--2891.

\bibitem[\citeproctext]{ref-ludecke2021performance}
Lüdecke, D., Ben-Shachar, M. S., Patil, I., Waggoner, P., \& Makowski,
D. (2021). Performance: An r package for assessment, comparison and
testing of statistical models. \emph{Journal of Open Source Software},
\emph{6}(60).

\bibitem[\citeproctext]{ref-luo2019robust}
Luo, Q., Rossion, B., \& Dzhelyova, M. (2019). A robust implicit measure
of facial attractiveness discrimination. \emph{Social Cognitive and
Affective Neuroscience}, \emph{14}(7), 737--746.

\bibitem[\citeproctext]{ref-makowski2019indices}
Makowski, D., Ben-Shachar, M. S., Chen, S. A., \& Lüdecke, D. (2019).
Indices of effect existence and significance in the bayesian framework.
\emph{Frontiers in Psychology}, \emph{10}, 2767.

\bibitem[\citeproctext]{ref-makowski2025too}
Makowski, D., Te, A. S., Neves, A., Kirk, S., Liang, N. Z., Mavros, P.,
\& Chen, S. A. (2025). Too beautiful to be fake: Attractive faces are
less likely to be judged as artificially generated. \emph{Acta
Psychologica}, \emph{252}, 104670.

\bibitem[\citeproctext]{ref-makowski2023we}
Makowski, D., \& Waggoner, P. D. (2023). Where are we going with
statistical computing? From mathematical statistics to collaborative
data science. In \emph{Mathematics} (8; Vol. 11, p. 1821). MDPI.

\bibitem[\citeproctext]{ref-monk2021beholding}
Monk Jr, E. P., Esposito, M. H., \& Lee, H. (2021). Beholding
inequality: Race, gender, and returns to physical attractiveness in the
united states. \emph{American Journal of Sociology}, \emph{127}(1),
194--241.

\bibitem[\citeproctext]{ref-ospina2012general}
Ospina, R., \& Ferrari, S. L. (2012). A general class of zero-or-one
inflated beta regression models. \emph{Computational Statistics \& Data
Analysis}, \emph{56}(6), 1609--1623.

\bibitem[\citeproctext]{ref-rouder2024hierarchical}
Rouder, J. N., \& Mehrvarz, M. (2024). Hierarchical-model insights for
planning and interpreting individual-difference studies of cognitive
abilities. \emph{Current Directions in Psychological Science},
\emph{33}(2), 128--135.

\bibitem[\citeproctext]{ref-vehtari2017practical}
Vehtari, A., Gelman, A., \& Gabry, J. (2017). Practical bayesian model
evaluation using leave-one-out cross-validation and WAIC.
\emph{Statistics and Computing}, \emph{27}, 1413--1432.

\bibitem[\citeproctext]{ref-williams2021beneath}
Williams, D. R., Mulder, J., Rouder, J. N., \& Rast, P. (2021). Beneath
the surface: Unearthing within-person variability and mean relations
with bayesian mixed models. \emph{Psychological Methods}, \emph{26}(1),
74.

\end{CSLReferences}

\end{document}