### 2025-06-12 analysis/2_study2.qmd
update

---
title: "Analysis - Study 2"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    toc: true
    code-fold: true
    self-contained: false
execute: 
  cache: true
  cache-lazy: false
---

```{r}
#| message: false
#| warning: false

options(mc.cores = parallel::detectCores() - 2)

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(ggdist)
library(brms)
library(cmdstanr)
library(rstan)
```

## Data Preprocessing

```{r}
df <- read.csv("https://raw.githubusercontent.com/RealityBending/FictionChoco/refs/heads/main/data/sample1.csv")
df$Sex <- factor(df$Sex, levels = c("Male", "Female"))
```

## Sample 1

### Attractiveness

```{r}
m <- lme4::lmer(Real ~ Sex / poly(Attractive, 2) + (1|Participant) + (1|Item), data = df)
display(parameters::parameters(m, effects = "fixed"))

estimate_relation(m, length = 30) |> 
  ggplot(aes(x=Attractive, y = Predicted)) + 
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high,, fill = Sex), alpha = 0.2) +
  geom_line(aes(color = Sex), linewidth = 1) +
  geom_smooth(data=df, aes(color = Sex, y = Real), method = "loess", se = FALSE, linetype = "dotted", formula = 'y ~ x') +
  scale_color_manual(values = c("Male"="#2196F3", "Female"="#E91E63")) +
  scale_fill_manual(values = c("Male"="#2196F3", "Female"="#E91E63")) +
  theme_minimal()
```

#### Model Description

::: {.panel-tabset}

##### Effect of Attractiveness

```{r}
m1_chocoat <- readRDS("models/sample1_chocoattractiveness.rds")

params_table <- function(m, keep = "polyAttractive") {
  params <- insight::find_parameters(m, effects = "fixed", flatten  = TRUE)
  d <- as.data.frame(m, variable = params[str_detect(params, keep)])
  tab <- bayestestR::describe_posterior(d, effects = "fixed", component = "all", 
                                        ci_method = "hdi", test = "p_direction")
  tab |> 
    mutate(Parameter = str_replace(Parameter, "b_Sex", "b_p_Sex"),
           Parameter = str_replace(Parameter, "b_Condition", "b_p_Condition")) |> 
    separate(Parameter, into = c("temp", "Parameter", "Effect"), sep = "_") |> 
    insight::format_table() |> 
    select(-temp) |> 
    mutate(col = tab$pd) |> 
    gt::gt() |> 
    gt::data_color(columns = "Parameter") |> 
    gt::data_color(columns = "col", method = "numeric", target_columns = "pd", 
                   row = col > 0.95, palette = "yellow") |> 
    gt::data_color(columns = "col", method = "numeric", target_columns = "pd", 
                   row = col > 0.97, palette = "green") |> 
    gt::cols_hide("col")
}
params_table(m1_chocoat, keep = "polyAttractive")
```

##### Random Effects

```{r}
table_dvour <- function(g) {
  performance::performance_dvour(g) |> 
    arrange(desc(D_vour)) |> 
    gt::gt() |> 
    gt::data_color("D_vour", palette = c("red", "red", "orange", "yellow", "green"), domain = c(0, 1)) |> 
    gt::fmt_auto()
}

g1 <- modelbased::estimate_grouplevel(m1_chocoat)
table_dvour(g1)
```

##### Full

```{r}
summary(m1_chocoat, robust = TRUE)
```

:::

#### Model Visualization

```{r}
visualize_dchoco <- function(m, newdata, x = c(-0.005, seq(0, 1, length.out = 1001), 1.005)) {

  p <- modelbased::estimate_prediction(m, data = newdata, predict = "mu", iterations = 100)
  confright <- modelbased::estimate_prediction(m, data = newdata, predict = "confright", iterations = 100)
  confleft <- modelbased::estimate_prediction(m, data = newdata, predict = "confleft", iterations = 100)
  precright <- modelbased::estimate_prediction(m, data = newdata, predict = "precright", iterations = 100)
  precleft <- modelbased::estimate_prediction(m, data = newdata, predict = "precleft", iterations = 100)
  if ("pex" %in% names(insight::find_parameters(m))) {
    pex <- modelbased::estimate_prediction(m, data = newdata, predict = "pex", iterations = 100)
  } else {
    pex <- list(Predicted = rep(0, nrow(newdata)))
  }
  if ("bex" %in% names(insight::find_parameters(m))) {
    bex <- modelbased::estimate_prediction(m, data = newdata, predict = "bex", iterations = 100)
  } else {
    bex <- list(Predicted = rep(0.5, nrow(newdata)))
  }
  if ("pmid" %in% names(insight::find_parameters(m))) {
    pmid <- modelbased::estimate_prediction(m, data = newdata, predict = "pmid", iterations = 100)
  } else {
    pmid <- list(Predicted = rep(0, nrow(newdata)))
  }

  densities <- data.frame()
  mus <- data.frame()
  for(i in 1:nrow(newdata)) {
    # Get distribution that will be shown
    d <- data.frame(
       x = x,
       y = cogmod::dchoco(x, p=p$Predicted[i], confright=confright$Predicted[i], confleft = confleft$Predicted[i],
                          precright = precright$Predicted[i], precleft = precleft$Predicted[i],
                          pex = pex$Predicted[i], bex = bex$Predicted[i], pmid = pmid$Predicted[i])
     )

    densities <- rbind(densities,  merge(d, newdata[i, ]))

    # Get mode
    x_mus <- c(-0.005, seq(0, 1, length.out = 1001), 1.005)
    d <- data.frame(
       x = x_mus,
       y = cogmod::dchoco(x_mus, p=p$Predicted[i], confright=confright$Predicted[i], confleft = confleft$Predicted[i],
                          precright = precright$Predicted[i], precleft = precleft$Predicted[i],
                          pex = pex$Predicted[i], bex = bex$Predicted[i], pmid = pmid$Predicted[i])
     )
    y <- c(max(d[d$x > 0.5, "y"]), max(d[d$x < 0.5, "y"]))
    mu <- data.frame(
     y = y ,
     x = c(d[d$x > 0.5 & d$y == y[1], "x"], 
           d[d$x < 0.5 & d$y == y[2], "x"]),
     side = c("confright", "confleft")
    )
    mus <- rbind(mus,  merge(mu, newdata[i, ]))
  }

  list(densities=densities, mu=mus)
}

pred1 <- visualize_dchoco(m1_chocoat,
                          newdata=insight::get_datagrid(m1_chocoat, length = 5, preserve_range = FALSE))

p3_density <- pred1$densities |> 
  mutate(Attractive = as.factor(Attractive),
         y = ifelse(x %in% c(0, 1), y * 3, y)) |>  # Inflate extremes for visibility
  ggplot(aes(x=x, y=y)) +
  geom_line(aes(color=Attractive), linewidth = 1.25) +
  geom_path(data = pred1$mu, aes(group = side), linewidth=0.3, lineend = 'round', 
            arrow = arrow(length = unit(0.03, "npc"), type = "closed")) +
  facet_wrap(~Sex) +
  labs(y = "Distribution", x = "Do you think this image is real?", color = "Attractiveness",
       title = "Effect on CHOCO distribution") +
  scale_colour_manual(values=scales::seq_gradient_pal("#FFCDD2", "#B71C1C", "Lab")(seq(0, 1, length.out=5))) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.25, 0.5, 0.75, 1), labels = c("      0%", "25%", "50%", "75%", "100%      ")) +
  # scale_linewidth_continuous(range = c(0.5, 2), guide = "none") +
  theme_minimal() +
  theme(axis.title.x = element_text(face = "bold"), 
        axis.text.y = element_blank(), 
        plot.title = element_text(face = "bold"),
        strip.background = element_rect(fill = "lightgrey", color = NA)) 
p3_density
```

```{r}
#| cache-lazy: false

colors <- c("p" = "#9C27B0", "confright" = "#795548", "confleft" = "#795548",
            "precright" = "#607D8B","precleft" = "#607D8B", "pex" = "#FF9800", "bex" = "#FF7811", "pmid" = "#4CAF50")

plot_params1 <- function(m, xlabel = "Attractiveness") {
  pred_params <- data.frame()
  for(param in c("mu", "confright", "confleft", "precright", "precleft", "pex", "bex")) {
    pred_params <- m |> 
      modelbased::estimate_prediction(data = "grid", length = 40, preserve_range = FALSE, 
                                      predict = param, iterations = 200) |>
      as.data.frame() |> 
      dplyr::mutate(Parameter = param) |> 
      rbind(pred_params)
  }

  if(xlabel == "Attractiveness") {
    pred_params$x <- pred_params$Attractive
  } else {
    pred_params$x <- pred_params$Beauty
  }

  pred_params |> 
    mutate(Parameter = ifelse(Parameter == "mu", "p", Parameter),
           Parameter = fct_relevel(Parameter, "p", "confright", "precright", "bex", "confleft", "precleft", "pex")) |> 
    filter(Parameter != "pex") |>
    ggplot(aes(x = x, y = Predicted)) +
    geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Parameter, group = Sex), alpha = 0.2) +
    geom_line(aes(color = Parameter, linetype = Sex), linewidth = 1) +
    facet_wrap(~Parameter, scales = "free_y", ncol=3, strip.position = "left") +
    labs(x = xlabel, title = "Effect on model parameters") +
    scale_fill_manual(values = colors, guide = "none") +
    scale_color_manual(values = colors, guide = "none") +
    scale_x_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1), labels = c(" 0", "0.25", "0.5", "0.75", "1 ")) +
    theme_minimal() +
    theme(strip.placement = "outside",
          axis.title.y = element_blank(),
          axis.title.x = element_text(face = "bold"),
          plot.title = element_text(face = "bold"),
          strip.text = element_text(size = rel(1), face = "italic"),
          panel.grid.minor.x = element_blank())
}

p3_params <- plot_params1(m1_chocoat, xlabel = "Attractiveness")

# Add significance
sig <- data.frame(
  Sex = c("Male", "Male", "Male", "Male", "Male"),
  Label = c("/", "U", "\\", "∩", "/"),
  Parameter = c("p", "confright", "confleft", "precleft", "bex"),
  x = 0.5,
  y = c(0.51, 0.6, 0.63, 3.75, 0.42)
) |> 
  mutate(Parameter =  fct_relevel(Parameter, "p", "confright", "bex", "confleft", "precleft"))

p3_paramssig <- p3_params + 
  geom_text(data = sig, aes(x = x, y = y, label = Label), 
            size = 5, color = "black", hjust = 0.5) 
p3_paramssig
```

### Beauty

```{r}
m <- lme4::lmer(Real ~ Sex / poly(Beauty, 2) + (1|Participant) + (1|Item), data = df)
display(parameters::parameters(m, effects = "fixed"))

estimate_relation(m, length = 30) |> 
  ggplot(aes(x=Beauty, y = Predicted)) + 
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high,, fill = Sex), alpha = 0.2) +
  geom_line(aes(color = Sex), linewidth = 1) +
  geom_smooth(data=df, aes(color = Sex, y = Real), method = "loess", se = FALSE, linetype = "dotted", formula = 'y ~ x') +
  scale_color_manual(values = c("Male"="#2196F3", "Female"="#E91E63")) +
  scale_fill_manual(values = c("Male"="#2196F3", "Female"="#E91E63")) +
  theme_minimal()
```

#### Model Description

::: {.panel-tabset}

##### Effect of Beauty

```{r}
m1_chocobeauty <- readRDS("models/sample1_chocobeauty.rds")
params_table(m1_chocobeauty, keep = "polyBeauty")
```

##### Random Effects

```{r}
table_dvour(modelbased::estimate_grouplevel(m1_chocobeauty))
```

##### Full

```{r}
summary(m1_chocobeauty, robust = TRUE)
```

:::

#### Model Visualization

```{r}
pred <- visualize_dchoco(m1_chocobeauty,
                         newdata=insight::get_datagrid(m1_chocobeauty, length = 5, preserve_range = FALSE))

p4_density <- pred$densities |> 
  mutate(Attractive = as.factor(Beauty),
         y = ifelse(x %in% c(0, 1), y * 3, y)) |>  # Inflate extremes for visibility
  ggplot(aes(x=x, y=y)) +
  geom_line(aes(color=Attractive), linewidth = 1.25) +
  # ggforce::geom_link2(data = pred$mu, aes(group = side, linewidth = Attractive, color = as.factor(Attractive)), 
  #                     lineend = 'round', n = 500) +
  geom_path(data = pred$mu, aes(group = side), linewidth=0, lineend = 'round', 
            arrow = arrow(length = unit(0.025, "npc"), type = "closed")) +
  facet_wrap(~Sex) +
  labs(y = "Distribution", x = "Do you think this image is real?", color = "Attractiveness") +
  scale_colour_manual(values=scales::seq_gradient_pal("#C5CAE9", "#283593", "Lab")(seq(0,1,length.out=5))) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.25, 0.5, 0.75, 1), labels = c("      0%", "25%", "50%", "75%", "100%      ")) +
  scale_linewidth_continuous(range = c(0.5, 2), guide = "none") +
  theme_minimal() +
  theme(axis.title.x = element_text(face = "bold"), 
        axis.text.y = element_blank(), 
        # axis.title.y = element_blank(), 
        strip.background = element_rect(fill = "lightgrey", color = NA)) 
p4_density
```

## Sample 2

```{r}
#| eval: false

set.seed(123)

dfraw <- read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace2/refs/heads/main/data/data_task.csv") |> 
  full_join(read.csv("https://raw.githubusercontent.com/RealityBending/FakeFace2/refs/heads/main/data/data_participants.csv") |> 
              select(Participant, Age), by = "Participant") |> 
  rename(Sex = Gender, Orientation = SexualOrientation, Item=Stimulus) |> 
  filter(Sex %in% c("Male", "Female"), Orientation %in% c("Heterosexual", "Bisexual")) |> 
  mutate(Relevance = case_when(
    Sex == "Male" & grepl("NF", Item) ~ "Relevant",
    Sex == "Female" & grepl("NM", Item) ~ "Relevant",
    .default = "Irrelevant"
  ))  |> 
  filter(Relevance == "Relevant") |> 
  mutate(Real = (Realness + 3) / 6,
         Attractive = Attractiveness / 6,
         Beauty = Beauty / 6,
         RealNoex = datawizard::change_scale(Real, to = c(0.125, 0.875)),
         SexOrientation = paste(Sex, "-", Orientation))

dfraw |> 
  select(Participant, Sex, Condition, Item, Real, RealNoex, Attractive, Beauty) |> 
  write.csv("../data/sample2.csv", row.names = FALSE)
```

```{r}
#| eval: false

# For demographic section
dfraw |> 
  mutate(n_stims = n(), .by = "Participant") |> 
  select(Participant, Age, Sex, n_stims) |> 
  slice(1, .by = "Participant") |> 
  report::report_participants()

report::report(dfraw$Sex)
```

```{r}
df <- read.csv("https://raw.githubusercontent.com/RealityBending/FictionChoco/refs/heads/main/data/sample2.csv")
df$Sex <- factor(df$Sex, levels = c("Male", "Female"))
df$Condition <- factor(df$Condition, levels = c("Photograph", "AI-Generated"))
```

### Attractiveness

```{r}
ggplot(df, aes(x=Real)) +
  geom_bar() +
  facet_wrap(~Sex)
```

```{r}
m <- lme4::lmer(RealNoex ~ Sex / poly(Attractive, 2) * Condition + (1|Participant) + (1|Item), data = df)
display(parameters::parameters(m, effects = "fixed"))

estimate_relation(m, length = 30) |> 
  ggplot(aes(x=Attractive, y = Predicted)) + 
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high,, fill = Condition), alpha = 0.2) +
  geom_line(aes(color = Condition), linewidth = 1) +
  geom_smooth(data=df, aes(color = Condition, y = RealNoex), method = "loess", se = FALSE, linetype = "dotted", formula = 'y ~ x') +
  scale_color_manual(values = c("AI-Generated"="#3F51B5", "Photograph"="#FF5722")) +
  scale_fill_manual(values = c("AI-Generated"="#3F51B5", "Photograph"="#FF5722")) +
  facet_wrap(~Sex) +
  theme_minimal()
```

#### Model Comparison

Interestingly, Beta-gate appears as a better fit, but this is likely related to the nature of the index used, WAIC, which heavily penalizes for the number of parameters (which is *much* larger for CHOCO). Probably not a very meaningful result, especially since the posterior checks clearly show its misfit.

```{r}
# m2_betagate <- readRDS("models/sample2_betagate.rds")
m2_betagate <- readRDS("models/sample2noex_betagate.rds")
# m2_chocoat <- readRDS("models/sample2_chocoattractiveness.rds")
m2_chocoat <- readRDS("models/sample2noex_chocoattractiveness.rds")
```

```{r}
loocomp <- loo::loo_compare(m2_betagate, m2_chocoat, criterion = "waic")

parameters(loocomp, include_ENP = TRUE)

report::report(loocomp)
```

##### Posterior Predictive Check

```{r}
make_ppcheck  <- function(m, fill = "#FF9800") {
  pred <- m |> 
    modelbased::estimate_prediction(df, keep_iterations = 100, iterations = 100) |>
    select(Sex, starts_with("iter_")) |> 
    bayestestR::reshape_iterations()

  pred |>
    ggplot(aes(x = iter_value)) +
    stat_bin(data = df, aes(x = RealNoex, y = after_stat(density)),
             geom="bar", breaks = seq(0, 1.0, by = 0.04), fill = "#424242") +
    stat_bin(aes(group = iter_group, y = after_stat(density)), position = "identity",
             geom="bar", breaks = seq(0, 1.0, by = 0.04), pad = TRUE,
             alpha = 0.01, fill = fill) +

    scale_y_sqrt(expand = c(0, 0.01)) +
    # scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.5, 1), labels = c("      0%", "50%", "100%      ")) +
    theme_minimal() +
    theme(axis.title.x = element_text(face = "bold"), 
          axis.title.y = element_blank(), 
          axis.text.y = element_blank(), 
          strip.background = element_rect(fill = "lightgrey", color = NA)) +
    facet_wrap(~Sex, scales = "free_y") +
    coord_cartesian(xlim = c(-0.003, 1.005), expand = TRUE)
}

p_ppc1 <- make_ppcheck(m2_chocoat, fill = "brown") +
  labs("Real") 
p_ppc2 <- make_ppcheck(m2_betagate, fill = "#FF9800") +
  labs("Real")

p_ppc1 / p_ppc2 
```

#### Model Description

::: {.panel-tabset}

##### Effect of Attractiveness

```{r}
params_table(m2_chocoat, keep = "polyAttractive|Condition")
```

##### Effect of Condition

```{r}
c <- data.frame()
for(param in c("mu", "confright", "confleft", "precright", "precleft", "pmid")) {
  c <- estimate_contrasts(m2_chocoat, contrast = "Condition", by = "Sex", predict = param, 
                          backend = "emmeans", test = "pd") |>
    mutate(Parameter = ifelse(param == "mu", "p", param)) |> 
    rbind(c)
}
c
```

```{r}
estimate_contrasts(m2_chocoat, contrast = "Condition", by = c("Attractive", "Sex"), 
                   predict = "mu", length = 5, backend = "emmeans", test = "pd") 
```

##### Full

```{r}
summary(m2_chocoat, robust = TRUE)
```

:::

#### Model Visualization

```{r}
pred2 <- visualize_dchoco(m2_chocoat,
                          newdata=insight::get_datagrid(m2_chocoat, length = 5, preserve_range = FALSE))

p4_density <- pred2$densities |> 
  mutate(Attractive = as.factor(Attractive)) |> 
         # y = ifelse(x %in% c(0, 1), y * 3, y),  # Inflate extremes for visibility
         # x = ifelse(Condition == "AI-Generated", x + 0.01, x)) |>  
  # filter(Condition == "Photograph") |>
  ggplot(aes(x=x, y=y)) +
  geom_line(aes(color=Attractive, linetype = Condition, linewidth = Condition)) +
  # geom_path(data = filter(pred2$mu, Condition == "Photograph"), aes(group = side), linewidth=0, lineend = 'round', 
  #           arrow = arrow(length = unit(0.025, "npc"), type = "closed")) +
  facet_wrap(~Sex) +
  labs(y = "Distribution", x = "Do you think this image is real?", color = "Attractiveness",
       title = "Effect on CHOCO distribution") +
  scale_colour_manual(values=scales::seq_gradient_pal("#FFCDD2", "#B71C1C", "Lab")(seq(0,1,length.out=5))) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.25, 0.5, 0.75, 1), labels = c("      0%", "25%", "50%", "75%", "100%      ")) +
  scale_linetype_manual(values = c("Photograph" = "solid", "AI-Generated" = "dotted")) +
  scale_linewidth_manual(values = c("Photograph" = 1.25, "AI-Generated" = 0.8)) +
  theme_minimal() +
  theme(axis.title.x = element_text(face = "bold"), 
        axis.text.y = element_blank(),  
        plot.title = element_text(face = "bold"),
        # axis.title.y = element_blank(), 
        strip.background = element_rect(fill = "lightgrey", color = NA)) 
p4_density
```

```{r}
# pred <- visualize_dchoco(m2_chocoat,
#                          newdata=insight::get_datagrid(m2_chocoat, length = 5, preserve_range = FALSE),
#                          x = seq(0, 1, length.out=7))
```

```{r}
#| cache-lazy: false

plot_params2 <- function(m, xlabel = "Attractiveness") {
  pred_params <- data.frame()
  for(param in c("mu", "confright", "confleft", "precright", "precleft", "pmid")) {
    pred_params <- m |> 
      modelbased::estimate_prediction(data = "grid", length = 40, preserve_range = FALSE, predict = param, iterations = 200) |>
      as.data.frame() |> 
      dplyr::mutate(Parameter = param) |> 
      rbind(pred_params)
  }

  if(xlabel == "Attractiveness") {
    pred_params$x <- pred_params$Attractive
  } else {
    pred_params$x <- pred_params$Beauty
  }

  pred_params |> 
    mutate(Parameter = ifelse(Parameter == "mu", "p", Parameter),
           Parameter = fct_relevel(Parameter, "p", "confright", "precright", "pmid", "confleft", "precleft")) |> 
    ggplot(aes(x = x, y = Predicted)) +
    geom_ribbon(aes(ymin = CI_low, ymax = CI_high, fill = Parameter, 
                    group = interaction(Sex, Condition), alpha = Condition)) +
    geom_line(aes(color = Parameter, linetype = Sex, 
                  alpha = interaction(Sex, Condition), group = interaction(Sex, Condition)), linewidth = 1) +
    facet_wrap(~Parameter, scales = "free_y", ncol=3, strip.position = "left") +
    labs(x = xlabel, title = "Effect on model parameters") +
    scale_fill_manual(values = colors, guide = "none") +
    scale_color_manual(values = colors, guide = "none") +
    scale_alpha_manual(values = c("Photograph" = 0.2, "AI-Generated" = 0, 
                                  "Male.Photograph" = 1, "Female.Photograph" = 1,
                                  "Male.AI-Generated" = 0.2, "Female.AI-Generated"=0.2), guide = "none") +
    scale_x_continuous(breaks = c(0, 0.25, 0.5, 0.75, 1), labels = c(" 0", "0.25", "0.5", "0.75", "1 ")) +
    theme_minimal() +
    theme(strip.placement = "outside",
          axis.title.y = element_blank(),
          axis.title.x = element_text(face = "bold"),
          plot.title = element_text(face = "bold"),
          strip.text = element_text(size = rel(1), face = "italic"),
          panel.grid.minor.x = element_blank())
}

p4_params <- plot_params2(m2_chocoat, xlabel = "Attractiveness")

# Add significance
sig <- data.frame(
  Sex = c("Male", "Male", "Female", "Male"),
  Label = c("/", "/", "U", "\\"),
  Parameter = c("p", "confright", "confright", "confleft"),
  x = 0.5,
  y = c(0.48, 0.425, 0.49, 0.48)
) |> 
  mutate(Parameter =  fct_relevel(Parameter, "p", "confright", "confleft"))

p4_paramssig <- p4_params + 
  geom_text(data = sig, aes(x = x, y = y, label = Label), 
            size = 5, color = "black", hjust = 0.5) 
p4_paramssig
```

### Beauty

```{r}
m <- lme4::lmer(RealNoex ~ Sex / poly(Beauty, 2) * Condition + (1|Participant) + (1|Item), data = df)
display(parameters::parameters(m, effects = "fixed"))

estimate_relation(m, length = 30) |> 
  ggplot(aes(x=Beauty, y = Predicted)) + 
  geom_ribbon(aes(ymin = CI_low, ymax = CI_high,, fill = Condition), alpha = 0.2) +
  geom_line(aes(color = Condition), linewidth = 1) +
  geom_smooth(data=df, aes(color = Condition, y = Real), method = "loess", se = FALSE, linetype = "dotted", formula = 'y ~ x') +
  scale_color_manual(values = c("AI-Generated"="#3F51B5", "Photograph"="#FF5722")) +
  scale_fill_manual(values = c("AI-Generated"="#3F51B5", "Photograph"="#FF5722")) +
  facet_wrap(~Sex) +
  theme_minimal()
```

#### Model Description

::: {.panel-tabset}

##### Effect of Beauty

```{r}
m2_chocobeauty <- readRDS("models/sample2noex_chocobeauty.rds")
params_table(m2_chocobeauty, keep = "polyBeauty|Condition")
```

##### Random Effects

```{r}
table_dvour(modelbased::estimate_grouplevel(m2_chocobeauty))
```

##### Full

```{r}
summary(m2_chocobeauty, robust = TRUE)
```

:::

#### Model Visualization

```{r}
pred <- visualize_dchoco(m2_chocobeauty,
                         newdata=insight::get_datagrid(m2_chocobeauty, length = 5, preserve_range = FALSE))

p5_density <- pred$densities |> 
  mutate(Beauty = as.factor(Beauty)) |> 
         # y = ifelse(x %in% c(0, 1), y * 3, y),  # Inflate extremes for visibility
         # x = ifelse(Condition == "AI-Generated", x + 0.01, x)) |>  
  # filter(Condition == "Photograph") |>
  ggplot(aes(x=x, y=y)) +
  geom_line(aes(color=Beauty, linetype = Condition, linewidth = Condition)) +
  # geom_path(data = filter(pred$mu, Condition == "Photograph"), aes(group = side), linewidth=0, lineend = 'round', 
  #           arrow = arrow(length = unit(0.025, "npc"), type = "closed")) +
  facet_wrap(~Sex) +
  labs(y = "Distribution", x = "Do you think this image is real?", color = "Attractiveness",
       title = "Effect on CHOCO distribution") +
  scale_colour_manual(values=scales::seq_gradient_pal("#C5CAE9", "#283593", "Lab")(seq(0,1,length.out=5))) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0, 0.25, 0.5, 0.75, 1), labels = c("      0%", "25%", "50%", "75%", "100%      ")) +
  scale_linetype_manual(values = c("Photograph" = "solid", "AI-Generated" = "dotted")) +
  scale_linewidth_manual(values = c("Photograph" = 1.25, "AI-Generated" = 0.8)) +
  theme_minimal() +
  theme(axis.title.x = element_text(face = "bold"), 
        axis.text.y = element_blank(),  
        plot.title = element_text(face = "bold"),
        # axis.title.y = element_blank(), 
        strip.background = element_rect(fill = "lightgrey", color = NA)) 
p5_density
```

## Figures

```{r}
#| eval: true
#| cache-lazy: false

p3_a <- wrap_elements((p3_density + theme(legend.position = "none")) / 
                        (p3_paramssig + theme(legend.position = "none")) + 
         patchwork::plot_annotation(title = "Sample 1", 
                                    theme = theme(plot.title = element_text(face = "bold.italic", hjust = 0.5))))

p4_a <- wrap_elements((p4_density + theme(axis.title.y = element_blank())) / 
                        p4_paramssig + 
         patchwork::plot_annotation(title = "Sample 2", 
                                    theme = theme(plot.title = element_text(face = "bold.italic", hjust = 0.5))))

p5 <- (p3_a | p4_a) + 
  plot_layout(guides = "collect", widths = c(0.45, 0.55))

ggsave("../paper/figures/fig5.png", p5, width = 13, height = 11, dpi = 300)
```