### 2025-06-21 analysis/0_proof.qmd
update based on comments

---
title: "Proof of Concept"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    toc: true
    code-fold: true
    self-contained: false
execute: 
  cache: true
  cache-lazy: false
---

```{r}
#| message: false
#| warning: false

options(mc.cores = parallel::detectCores() - 2)

library(tidyverse)
library(easystats)
library(patchwork)
library(ggside)
library(ggdist)
library(brms)
library(cmdstanr)
library(rstan)
library(cogmod)
library(gt)
library(webshot2)
Sys.setenv(CHROMOTE_CHROME = "C:/Program Files (x86)/Microsoft/Edge/Application/msedge.exe")
```

## Data

```{r}
set.seed(33)

c <- c("Condition A: CHOCO(p = 0.5, confleft = 0.3, confright = 0.3)",
       "Condition B: CHOCO(p = 0.3, confleft = 0.3, confright = 0.7)")

df <- data.frame(
  y = c(
    rchoco(1000, p = 0.5, confleft = 0.3, confright = 0.3, pex = 0),
    rchoco(1000, p = 0.3, confleft = 0.3, confright = 0.7, pex = 0)
    ),
  Condition = c(rep(c("A", "B"), each = 1000))
)

p1 <- data.frame(
  x = c(seq(0, 1, length.out = 1000), seq(0, 1, length.out = 1000)),
  y = c(
    dchoco(seq(0, 1, length.out = 1000), p = 0.5, confleft = 0.3, confright = 0.3, pex = 0),
    dchoco(seq(0, 1, length.out = 1000), p = 0.3, confleft = 0.3, confright = 0.7, pex = 0)
  ),
  Condition = rep(c, each = 1000)
) |> 
  ggplot(aes(x = x, y = y)) +
  geom_histogram(data = df, aes(x = y, y = after_stat(density), fill = Condition),
                bins = 35, alpha = 0.5, position = "identity") +
  geom_line(aes(color = Condition), linewidth = 2, alpha = 0.95) +
  scale_fill_manual(values = c("#0072B2", "#FF5722"), guide = "none") +
  scale_color_manual(values = c("#0072B2", "#FF5722")) +
  labs(x = "Confidence", y = "Distribution", title = "Usage Example") +
  theme_minimal() +
  theme(legend.position = "top",
        legend.direction='vertical',
        legend.title = element_blank(),
        legend.justification = c(0, 0.5),
        axis.text.y = element_blank(),
        plot.title = element_text(face = "bold"))
p1
```

## Linear model

```{r}

model1 <- lm(y ~ Condition, data = df)

t1 <- parameters::parameters(model1) |> 
  format_table(zap_small = TRUE) |> 
  data_select(exclude = "SE") |> 
  data_rename(select = "Parameter", "Effect") |> 
  gt() |> 
  data_color(
    columns = "Effect",
    palette = c("#0072B2", "#FF5722")) |>
  data_color(
    columns = "p",
    rows = 2,
    palette = "red") |> 
  tab_header(
    title = html("<strong>Linear Model</strong>"),
    subtitle = html("<p style='text-align: left;'><code>lm(y ~ Condition)</code><br></p>")) |> 
  cols_align(columns = -Effect, align = "right")|>
  cols_label(Effect = md("***Effect***"),
             `95% CI` = md("***95% CI***"),
             Coefficient = md("***Coefficient***"),
             `t(1998)` = md("***t(1998)***"),
             p = md("***p***")) 
t1

gtsave(t1, "table1.png", path="./", vwidth = 1280)
p_t1 <- png::readPNG("table1.png") |> 
  grid::rasterGrob()
file.remove("table1.png")
```

## CHOCO

```{r}
model2 <- brm(bf(
    y ~ Condition,
    confleft ~ Condition,
    confright ~ Condition,
    precright ~ 1,
    precleft ~ 1,
    pex = 0,
    bex = 0.5,
    pmid = 0),
  data = df,
  family = choco(link_mu = "identity", 
                 link_confleft = "identity", 
                 link_confright = "identity"),
  stanvars = cogmod::choco_stanvars(),
  init = 0.5,
  iter = 2000,
  warmup = 1000,
  chains = 4,
  refresh = 0,
  cores = parallel::detectCores() - 2,
  backend = "cmdstanr"
)

t2 <- parameters::parameters(model2, dispersion = TRUE) |> 
  as.data.frame() |> 
  mutate(t = Median / MAD, p = bayestestR::pd_to_p(pd)) |> 
  head(6) |> 
  select(Parameter = Component, Effect = Parameter, Coefficient = Median, CI, CI_low, CI_high, t, p) |> 
  mutate(Effect = rep(c("(Intercept)", "Condition [B]"), 3),
         Parameter = ifelse(Parameter == "conditional", "p", Parameter)) |> 
  format_table(zap_small = TRUE) |> 
  gt() |> 
  data_color(
    columns = "Effect",
    palette = rep(c("#0072B2", "#FF5722"), 3)) |> 
  data_color(
    columns = "Parameter",
    palette = c("#9a7b78", "#c09569", "#9C27B0")) |> 
  data_color(
    columns = "p",
    rows = c(2, 6),
    palette = "green") |>
   data_color(
    columns = "p",
    rows = c(4),
    palette = "red") |>
  tab_header(
    title = html("<strong>CHOCO Model</strong>"),
    subtitle = html("<p style='text-align: left;'><code>brm(bf(<br>&emsp;&emsp;y ~ Condition, <br>&emsp;&emsp;confleft ~ Condition, <br>&emsp;&emsp;confright ~ Condition), <br>family = cogmod::choco())</code><br></p>")) |>
  cols_align(columns = -c(Effect, Parameter), align = "right") |>
  cols_label(Effect = md("***Effect***"),
             Parameter = md("***Parameter***"),
             Coefficient = md("***Coefficient***"),
             `95% CI` = md("***95% CI***"),
             `t` = md("***t***"),
             `p` = md("***p***"))

t2

gtsave(t2, "table2.png", path="./", vwidth = 1280)
p_t2 <- png::readPNG("table2.png") |> 
  grid::rasterGrob()
file.remove("table2.png")
```

## Figure

```{r}
# p2 <- wrap_elements(wrap_elements(p_t1) / wrap_elements(p_t2) + 
#          patchwork::plot_annotation(title = "Effect of Condition", theme = theme(plot.title = element_text(face = "bold"))) +
#          patchwork::plot_layout(heights = c(0.35, 0.65)))
p2 <- gridExtra::grid.arrange(p_t1, p_t2, heights = c(0.35, 0.65))

ggsave("table3.png", p2, width = 10, height = 10, dpi = 300)
p2 <- png::readPNG("table3.png") |> 
  grid::rasterGrob()
file.remove("table3.png")

fig <- gridExtra::grid.arrange(p1, p2, ncol = 2, widths = c(0.45, 0.55))
ggsave("../paper/figures/fig2b.png", fig, dpi = 300, width = 9, height=6)

# ----
fig2a <- grid::rasterGrob(png::readPNG("../paper/figures/fig2.png"))
fig2b <- grid::rasterGrob(png::readPNG("../paper/figures/fig2b.png"))

figfinal <- gridExtra::grid.arrange(fig2a, fig2b,
  ncol = 1, heights = c(0.5, 0.5))
figfinal
ggsave("../paper/figures/fig2final.png", figfinal, dpi = 600, 
       width = grid::resolveRasterSize(fig2a)$width,
       height = grid::resolveRasterSize(fig2a)$height + grid::resolveRasterSize(fig2b)$height)
```